---
title: "DATA 605 - Assignment 15"
author: "Amber Ferger"
date: "5/11/2020"
output: html_document
---

## Problem 1
**Find the equation of the regression line for the given points. Round any final values to the nearest hundredth, if necessary.**  
$$(5.6, 8.8), (6.3, 12.4), (7, 14.8), (7.7, 18.2), (8.4, 20.8)$$

## Solution
The regression line can be modeled as $y=mx+b$. Given the points above, we can solve for $m$ and $b$ by using the following equations:  

$$m = \frac{n \sum_{i=1}^{n} x_i y_i - \sum_{i=1}^{n} x_i \sum_{i=1}^{n}y_i}{n \sum_{i=1}^{n}x_i^2 - (\sum_{i=1}^{n}x_i)^2}$$

$$b = \frac{\sum_{i=1}^{n}y_i - m \sum_{i=1}^{n} x_i}{n}$$

```{r}

df <- data.frame("x" = c(5.6,6.3,7,7.7,8.4), "y" = c(8.8,12.4,14.8,18.2,20.8))
df$xy <- df$x*df$y
df$x2 <- df$x * df$x

n <- nrow(df)
sumx <- sum(df$x)
sumy <- sum(df$y)
sumxy <- sum(df$xy)
sumx2 <- sum(df$x2)

m <- ((n * sumxy) - (sumx * sumy))/((n * sumx2) - (sumx)^2)
b <- (sumy - (m*sumx) )/ n

round(m,2)
round(b,2)

```

Therefore, we know our final equation is: 
$$y = 4.26x-14.8$$

## Problem 2
**Find all local maxima, local minima, and saddle points for the function given below. Write your answer(s) in the form** $(x,y,z)$**. Separate multiple points with a comma.**
$$f(x,y) = 24x - 6xy^2 - 8y^3$$

## Solution
First, identify the derivatives with respect to $x$ and $y$:
$$F_x(x,y) = 24 - 6y^2$$
$$F_y(x,y) = -12xy-24y^2$$

Find the critical points of $y$ by setting $F_x(x,y) = 0$ and solving for $y$. 
$$F_x(x,y) = 24 - 6y^2$$
$$6y^2 = 24 $$
$$y^2 = 4 $$
$$y=2, y=-2$$
Find the critical points of $x$ by subbing $y=2$ and $y=-2$ into $F_y(x,y)$:
$$F_y(x,y) = -12xy-24y^2$$
For $y=2$:
$$F_y(x,2)= -12x(2)-24(2^2)$$
$$24x=-96$$
$$x=-4$$
For $y=-2$:
$$F_y(x,-2)= -12x(-2)-24(-2^2)$$
$$-24x=-96$$
$$x=4$$

Another critical point is $(0,0)$ because this also solves $F_y(x,y)$:
$$F_y(x,y) = -12xy-24y^2$$
$$ 12xy= -24y^2$$
$$ xy= -2y^2$$
$$ x= -2y$$

Next, using $F_x(x,y)$ and $F_y(x,y)$, find the second derivatives in terms of $F_{xx}(x,y)$, $F_{yy}(x,y)$, and $F_{xy}(x,y)$:

$$F_{xx}(x,y) = 0$$
$$F_{yy}(x,y) = -12x - 48y$$
$$F_{xy}(x,y) = -12y$$
Use the second derivative test provided the rules defined:  
$D = D(x,y) = F_{xx}(x,y) \cdot F_{yy}(x,y) - (F_{xy}(x,y))^2$  

$D >0$ and $F_{xx}(x,y) > 0$, then $F(x,y)$ is a local minimum.  

$D >0$ and $F_{xx}(x,y) < 0$, then $F(x,y)$ is a local maximum.

$D < 0$ then $F(x,y)$ is a saddle point.   


The critical points are $(-4,2)$ , $(4,-2)$, and $(0,0)$ and will be used to solve for $D$:  
At $(-4,2)$:  
```{r}
x <- -4
y <- 2
fxx <- 0
fyy <- (-12*x) - (48*y)
fxy <- -12*y

d <- (fxx * fyy) - (fxy)^2
d

```

At $(4,-2)$: 

```{r}
x <- 4
y <- -2
fxx <- 0
fyy <- (-12*x) - (48*y)
fxy <- -12*y
d <- (fxx * fyy) - (fxy)^2

d

```

At $(0,0)$: 

```{r}
x <- 0
y <- 0
fxx <- 0
fyy <- (-12*x) - (48*y)
fxy <- -12*y
d <- (fxx * fyy) - (fxy)^2
d

```

This tells us that we have 2 saddle points: $(-4,2)$ and $(4,-2)$. We cannot make a conclusion about $(0,0)$ using the second derivatives test.  

```{r}
x <- -4
y <- 2
z <- 24*x−6*x*(y^2)−8*(y^3)
z

x <- 4
y <- -2
z <- 24*x−6*x*(y^2)−8*(y^3)
z  
```
**Final solution:**  
$$(-4,2,-64),(4,-2,64)$$

## Problem 3
**A grocery store sells two brands of a product, the "house" brand and a "name" brand. The manager estimates that if she sells the "house" brand for x dollars and the "name" brand for y dollars, she will be able to sell $81 - 21x + 17y$ units of the "house" brand and $40 + 11x - 23y$ units of the "name" brand.**  
  
**Step 1. Find the revenue function** $R(x,y)$.  
  
**Step 2. What is the revenue if she sells the "house" brand for \$2.30 and the "name" brand for \$4.10?**

## Solution

Revenue = Number of house brand units * cost of house brand + Number of name brand units * cost of name brand  

$$R(x,y) = (81 - 21x + 17y)x + (40 + 11x - 23y)y$$
$$R(x,y) = 81x - 21x^2 + 17yx + 40y + 11xy - 23y^2$$
$$R(x,y) = 81x - 21x^2 + 28yx + 40y - 23y^2$$


```{r}

x <- 2.30
y <- 4.10
r <- 81*x - 21*x^2 + 28*y*x + 40*y - 23*y^2
r

```

## Problem 4
**A company has a plant in Los Angeles and a plant in Denver. The firm is committed to produce a total of** $96$ **units of a product each week. The total weekly cost is given by** $C(x, y) = \frac{1}{6}x^2 + \frac{1}{6}y^2 + 7x + 25y + 700$**, where** $x$ **is the number of units produced in Los Angeles and** $y$ **is the number of units produced in Denver. How many units should be produced in each plant to minimize the total weekly cost?**

## Solution
From the problem statement, we know that $x+y = 96$, or $x = 96-y$. We will use this to rewrite our function: 
$$C(x, y) = \frac{1}{6}x^2 + \frac{1}{6}y^2 + 7x + 25y + 700$$
$$C(y) = \frac{1}{6}(96-y)^2 + \frac{1}{6}y^2 + 7(96-y) + 25y + 700$$
$$C(y) = \frac{1}{6}(y^2-192y+9216) + \frac{1}{6}y^2 - 7y+ 672 + 25y + 700$$

$$C(y) = \frac{1}{6}y^2- 32y +1536 + \frac{1}{6}y^2 +18y + 1372$$

$$C(y) = \frac{1}{3}y^2 - 14y + 2908$$

To find the minimum value, we will take the derivative of $C(y)$ and set equal to $0$ to solve for $y$.

$$C'(y) = \frac{2}{3}y - 14$$
$$0 = \frac{2}{3}y - 14$$
$$\frac{2}{3}y = 14$$
$$y = 21$$

We can plug this back into our original equation $x = 96-y$ to get the value of $x$:

$$x = 96-21 = 75$$

Our final values are: $(75, 21)$. 


## Problem 5
**Evaluate the double integral on the given region.**
$$\int \int_{R}(e^{8x+3y})dA$$
$R:2 \leq x \leq 4$ and $2 \leq y \leq 4$  
**Write your answer in exact form without decimals.**

$$\int_2^4 \int_2^4(e^{8x+3y})dxdy$$
$$\frac{1}{8} \int_2^4 e^{8x+3y}\Big|_2^4 dy$$
$$\frac{1}{8} \int_2^4 (e^{32+3y} - e^{16+3y}) dy$$
$$\frac{1}{8} \cdot \frac{1}{3} \Big(\int_2^4 e^{32+3y}dy - \int_2^4e^{16+3y} dy\Big)$$
$$\frac{1}{24}  \Big(e^{32+3y}\Big|_2^4 - e^{16+3y}\Big |_2^4 \Big)$$
$$\frac{1}{24}  \Big(e^{44} - e^{38} - e^{28} - e^{22} \Big)$$

