---
title: "DATA 605 - Assignment 7"
author: "Amber Ferger"
date: "3/14/2020"
output: html_document
---

## Problem 1  
**Let $X_1$, $X_2$...$X_n$ be $n$ mutually independent random variables, each of which is uniformly distributed on the integers from 1 to $k$. Let $Y$ denote the minimum of the $X_i$s. Find the distribution of $Y$.**  

Let $j$ be a value within the range $[1,k]$.

We will define the probability of $Y=j$ as: 
$$P(Y=j) = \frac{num\; ways \: event \; can \; happen}{total \; number \; of \; options}$$

We know that the total number of options is $k^n$ because each of the $n$ variables can take on $k$ different values. Therefore, $k^n$ will form the denominator in our above equation.  

Intuitively, we know that as $j$ increases, the probability of the minimum of the $X_is$ being equal to $j$ decreases. This is because the number of ways that $j$ can be the minimum of the set of variables (our numerator in the probability equation above) decreases. In order to abstract this, let's use an example.  

Suppose $k=4$ and $n=2$. In order for $Y=j$ to be true, we know that the set of values each variable can take on is as follows:  

$Y=j=1: \{1,2,3,4\}$  
$Y=j=2: \{2,3,4\}$  
$Y=j=3: \{3,4\}$  
$Y=j=4: \{4\}$  

This means that in order for the minimum value to be 3, the variables cannot take on a value of 1 or 2. Similarly, in order for the minimum value to be 4, the variables can only take on a value of 4. 

The number of ways the event can occur for each value of $j$ can be modeled as:  
$$number \; of \; values \; in \; the \;set^{number \; of \; variables}$$  
Following our example:  

$Y=j=1: 4^2$  
$Y=j=2: 3^2$  
$Y=j=3: 2^2$  
$Y=j=4: 1^2$  

We can abstract the number of times Y=j can occur as:  
$$Y=j: (k-j+1)^n$$
Putting this all together:  

$$P(Y=j) = \frac{(k-j+1)^n}{k^n}$$

## Problem 2 
**Your organization owns a copier (future lawyers, etc.) or MRI (future doctors). This machine has a manufacturer's expected lifetime of 10 years. This means that we expect one failure every ten years. (Include the probability statements and R code for each part).**  
We know that a "success" is the probability that the event will happen. In this instance, success is defined as the probability that a machine will fail. A failure is defined as the probability that the machine will last. For all of the following, let:  
$$P(Success)=p = \frac{1}{10}$$
$$P(Failure)=1-p = \frac{9}{10}$$ 
$$X=number\; of\; trials$$ 

**a. What is the probability that the machine will fail after 8 years? Provide also the expected value and standard deviation. Model as a geometric. (Hint: the probability is equivalent to not failing during the first 8 years.)**  
Modeling as a geometric distribution, we know that: 
$$P(X>k)=(1-p)^k$$

where $k$ is the total number of years.  

Therefore, the probability of 0 successes (machine failures) in the first 8 years is: 
```{r}
p <- 1/10
k <- 8

geomDist <- (1-p)^k
geomDist

```

The expected value is: 
$$E=\frac{1}{p}=10$$

The standard deviation is: 
$$\sigma = \sqrt{\frac{1-p}{p^2}}$$

```{r}

geomStDev <- sqrt((1-p)/(p)^2)
geomStDev

```

**b. What is the probability that the machine will fail after 8 years? Provide also the expected value and standard deviation. Model as an exponential.**  

Modeling as an exponential distribution, we know that:   
$$P(X\leq k) = 1-e^{-k\lambda}$$

where $k$ is the total number of years and $\lambda$ is the average number of failures per year.  

The complement of this is: 
$$P(X> k)=e^{-k\lambda}$$

Therefore, the probability of 0 successes (machine failures) in the first 8 years is: 
```{r}

k <- 8
expLambda <- 1/10
expDist <- exp(-k*expLambda)
expDist

```

The expected value is: 
$$E = \frac{1}{\lambda}$$
```{r}

expExpVal <- 1/expLambda
expExpVal

```

The standard deviation is:  
$$\sigma = \sqrt{\frac{1}{\lambda^2}}$$
```{r}

stDevExp <- sqrt(1/(expLambda^2))
stDevExp

```

**c. What is the probability that the machine will fail after 8 years? Provide also the expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8 years)**  
The binomial distribution is: 
$$b(n,p,k) = {{n}\choose{k}}p^k(1-p)^{n-k}$$

where $n$ is the number of years  and $k$ is the number of successes. 

Therefore, the probability of 0 successes (machine failures) in the first 8 years is: 
```{r}

n <- 8
k <- 0
p <- 1/10
choose(n,k)*p^k*(1-p)^(n-k)


```

The expected value is:  

```{r}

expBinom <- n * p
expBinom

```

The standard deviation is: 
```{r}
stDevBinom <- sqrt(n*p*(1-p))
stDevBinom

```

**d. What is the probability that the machine will fail after 8 years? Provide also the expected value and standard deviation. Model as a Poisson.** 

We know that the mean failure over the 8 year period is: 
$$\lambda = n * p$$
```{r}

poissonLambda <- n * p
poissonLambda

```

We can model this as a Poisson distribution: 

$$P(X=n)= \frac{\lambda^ne^{-\lambda}}{n!}$$

where $n$ is the number of failures in the first 8 years.

Therefore, the probability of 0 successes (machine failures) in the first 8 years is: 
```{r}

n <- 0
poissonDist <- ((poissonLambda^n)*exp(-poissonLambda))/factorial(n)

poissonDist

```

The expected value is:  
$$E(X) = \lambda = 0.8$$  

In a poisson distribution, we know that: 
$$E(X) = Var(X) = \lambda$$  

Since $Var(X) = \sqrt{\sigma}$, we know that the standard deviation is:   

```{r}
stDevPoisson <- sqrt(poissonLambda)
stDevPoisson


```
