---
title: "DATA 605 - Assignment 5"
author: "Amber Ferger"
date: "2/24/2020"
output: html_document
---

**Choose independently two numbers, $B$ and $C$, at random from the interval $[0,1]$ with uniform density. Prove that $B$ and $C$ are proper probability distributions. Note that the point $(B,C)$ is then chosen at random in the unit square.**  

First, we will define our interval values and simulate the draw of $B$ and $C$ from a uniform distriubution 10,000 times. 
``` {r}
set.seed(123)

# Set number of runs to 10,000
n <- 10000
minVal <- 0
maxVal <- 1


setB <- runif(n, min = minVal, max = maxVal)
setC <- runif(n, min = minVal, max = maxVal)

```

To prove that $B$ and $C$ are proper probability distributions, let's first make sure that all values are within the interval range [0,1].


``` {r}
# check to see if any values are below 0
paste0('Percent of times B is in correct range: ',100*length(setB[which(setB >0 & setB < 1)])/n)
paste0('Percent of times C is in correct range: ',100*length(setC[which(setC >0 & setC < 1)])/n)

```

Next, we will visually examine the distributions of $B$ and $C$ against a proper uniform distribution: 
```{r}
hist(setB, 
     freq = FALSE, 
     xlab = 'B',  
     density = 20,
     main = "Distribution of B")
curve(dunif(x, min = minVal, max = maxVal), 
      from =-0.50, to = 1.5, 
      n = n, 
      col = "darkblue", 
      lwd = 2, 
      add = TRUE, 
      yaxt = "n",
      ylab = 'probability')

```

``` {r}
hist(setC, 
     freq = FALSE, 
     xlab = 'B',  
     density = 20,
     main = "Distribution of C")
curve(dunif(x, min = minVal, max = maxVal), 
      from =-0.50, to = 1.5, 
      n = n, 
      col = "darkblue", 
      lwd = 2, 
      add = TRUE, 
      yaxt = "n",
      ylab = 'probability')

```

Both are nearly uniform distributions, so we can say that $B$ and $C$ follow proper probability distributions. 

**Find the probability that: **  

(a) $B + C < \frac{1}{2}$
``` {r}

sum((setB + setC)< 0.5)/n

```

(b) $BC < \frac{1}{2}$  
``` {r}

sum((setB * setC)< 0.5)/n

```


(c) $|B-C| < \frac{1}{2}$  
``` {r}
sum(abs((setB - setC))< 0.5)/n


```


(d) $max\{B,C\} < \frac{1}{2}$  
``` {r}

finalCount <- 0

for(i in seq(n)){
  finalCount <- finalCount +(max(setB[i],setC[i]) < 0.5)
}

finalCount/n

```


(e) $min\{B,C\} < \frac{1}{2}$  
``` {r}


finalCount <- 0

for(i in seq(n)){
  finalCount <- finalCount +(min(setB[i],setC[i]) < 0.5)
}

finalCount/n

```
