---
title: "DATA 605 - Assignment 2, Problem Set 2"
author: "Amber Ferger"
date: "2/9/2020"
output: html_document
---

# Problem Set 2
**Matrix factorization is a very important problem. There are supercomputers built just to do matrix factorizations. Every second you are on an airplane, matrices are being factorized. Radars that track flights use a technique called Kalman filtering. At the heart of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear systems of equations when they track your flight using radars. Write an R function to factorize a square matrix A into LU or LDU, whichever you
prefer.**

I am using the following pseudocode algorithm for LU decomposition: https://www.researchgate.net/figure/a-LU-decomposition-pseudocode-b-Example-on-LU-decomposition_fig4_220624910

``` {r}
luFunction <- function(A){
  
  # this method only words for square matrices 
  if (nrow(A) != ncol(A)){
    print('LU Decomposition only works with square matrices!')
    break
  }

  n <- nrow(A)
  
  # set up blank matrices for L and U
  L <- matrix(, nrow = n, ncol = n)
  U <- matrix(, nrow = n, ncol = n)


  for (k in 1:n){
    U[k,k] <- A[k,k]
    L[k,k] <- 1
    
    if (k+1 <= n){
      
      # pivot column calculations
      for (i in (k+1):n){
        L[i,k] <- A[i,k]/U[k,k]
        U[k,i] <- A[k,i]
      }
    
      # remaining matrix update
      for (i in (k+1):n){
      
        for (j in (k+1):n){
          A[i,j] <- A[i,j] - (L[i,k]%*%U[k,j])
        }
      }
      
    } # if statement

  }

  L[is.na(L)] <- 0
  U[is.na(U)] <- 0

  
  print('The L matrix is:')
  print(L)

  
  print('The U matrix is:')
  print(U)
#### function end  
}
  
```

Now that we have our function built, let's check it with a square matrix: 
``` {r}

A <- matrix(
  c(1, 1, 3, 2, -1, 5, -1, -2, 4),
  nrow = 3,
  ncol = 3,
  byrow = TRUE
)


luFunction(A)
```

