---
title: "DATA 605 - Assignment 4, Problem Set 1"
author: "Amber Ferger"
date: "2/17/2020"
output: html_document
---

# Problem Set 1
**In this problem, we'll verify using R that SVD and Eigenvalues are related as worked out in the weekly module. Given a 3 x 2 matrix A:**  

$$A = \left[\begin{array}{cc} 
1 & 2 & 3  \\
-1 & 0 & 4 
\end{array}\right]$$  

**write code in R to compute ** $X = AA^T$ **and** $Y = A^TA$. **Then, compute the eigenvalues and eigenvectors of X and Y using the built-in commands in R. Then compute the left-singular, singular values, and right-singular vectors of A using the *svd* command. Examine the two sets of singular vectors and show that they are indeed eigenvectors of X and Y. In addition, the two non-zero eigenvalues (the 3rd value will be very close to zero, if not zero) of both X and Y are the same and are squares of the non-zero singular values of A. **


## Solution  
Define $A$, $A^{T}$, $X$, and $Y$:
```{r}
A <- matrix(c(1,-1,2,0,3,4),nrow=2)
At <- t(A)

X <- A%*%At
Y <- At%*%A

```

Define the eigenvectors and eigenvalues of $X$ and $Y$:

```{r}
# eigenvalues and eigenvectors of X
evX <- eigen(X)
eValX <- evX$values
eVecX <- evX$vectors

# eigenvalues and eigenvectors of Y
evY <- eigen(Y)
eValY <- evY$values
eVecY <- evY$vectors
```

Compute the left-singular, singular values, and right-singular vectors of $A$:

``` {r}
svdA <- svd(A)
leftVals <- svdA$u
singVals <- svdA$d
rightVals <- svdA$v

```


Examine the two sets of singular vectors and show that they are indeed eigenvectors of $X$ and $Y$.  

Left singular values of $A$ are eigenvectors of $X$:

``` {r}
leftVals
eVecX
```

Right singular values of $A$ are eigenvectors of $Y$:
```{r}
rightVals
eVecY

```


The two non-zero eigenvalues (the 3rd value will be very close to zero, if not zero) of both $X$ and $Y$ are the same and are squares of the non-zero singular values of $A$.

``` {r}
singVals
eValX

singVals^2 - eValX

```

```{r}
singVals
eValY

singVals^2 - eValY[1:2]

```