---
title: "DATA 605 - Assignment 4, Problem Set 2"
author: "Amber Ferger"
date: "2/17/2020"
output: html_document
---

# Problem Set 2
**Using the procedure outlined in section 1 of the weekly handout, write a function to compute the inverse of a well-conditioned full-rank square matrix using co-factors. In order to compute the co-factors,  you may use built-in commands to compute the determinant. Your function should have the following signature: **  

$$B = myinverse(A)$$  

**where A is a matrix and B is its inverse and AxB = I. The off-diagonal elements of I should be close to zero, if not zero. Likewise, the diagonal elements should be close to 1, if not 1. Small numerical precision errors are acceptable but the function *myinverse* should be correct and must use co-factors and determinant of A to compute the inverse.**  

## Solution

First we'll define the function *myInverse*. The approach will be to cycle through our matrix and determine each element of the cofactor matrix, *C*. We will then compute $A^{-1} = \frac{C^T}{det(A)}$.

``` {r}
myInverse <- function(A){
  
  n <- nrow(A) # define the dimensions of the matrix
  dimList <- seq(nrow(A)) # create a sequence for the for loop to iterate over
  tempList <- matrix(, nrow = n, ncol = n) # store cofactors here
  
  # for all rows
  for(i in dimList){
  
    # set row indices
    if (i==1){
      rows <- seq((i+1),n)
    } else if (i==n) {
      rows <- seq(1,(i-1))
    } else {
      rowa <- seq(1,(i-1))
      rowb <- seq((i+1),n)
      rows <- c(rowa,rowb)
    }
    
    # for all columns
    for(j in dimList){
  
      # set col indices
      if (j==1){
        cols <- seq((j+1),n)
      } else if (j==n) {
        cols <- seq(1,(j-1))
      } else {
        cola <- seq(1,(j-1))
        colb <- seq((j+1),n)
        cols <- c(cola,colb)
      }
      
      if(n==2){
        finalVal <- ((-1)^(i+j)) * A[rows,cols]
      } else {
        cofacDet <- det(A[rows,cols])
        finalVal <- (-1^(i+j))* cofacDet
      }
      
      # append the cofactor to the appropriate element of the tempList matrix
      tempList[i,j] <- finalVal
    }
  }
  
  Ainv <- t(tempList)/det(A)
  
  return(Ainv)
  
}

```

We will test our function using the matrix:
$$
A = \left[\begin{array}{cc} 
1 & 2   \\
3 & 4 
\end{array}\right]
$$ 

First, let's generate the inverse function: 
``` {r}

A <- matrix(c(1,3,2,4),nrow=2)
myInv <- myInverse(A)
myInv

```

Next, we will verify that $AA^{-1}=I$: 

```{r}

A%*%myInv

```