---
title: "DATA 605 - Assignment 12"
author: "Amber Ferger"
date: "4/20/2020"
output: html_document
---

## Data Description

**The attached *who.csv* dataset contains real-world data from 2008. The variables included are: **  

* **$Country$: name of country**
* **$LifeExp$: average life expectancy for the country in years**
* **$InfantSurvival$: proportion of those surviving to one year or more**
* **$Under5Survival$: proportion of those surviving to five years or more**
* **$TBFree$: proportion of the population without TB**
* **$PropMD$: proportion of the population who are MDs**
* **$PropRN$: proportion of the population are RNs**
* **$PersExp$: mean personal expenditures on healthcare in US dollars at average exchange rate**
* **$GovtExp$: mean government expenditures per capita on healthcare, US dollars at average exchange rate**
* **$TotExp$: sum of personal and government expenditures**

```{r}
library(readr)

who.df <- read_csv('https://raw.githubusercontent.com/amberferger/DATA605_Homework/master/AFerger_Assignment12_Data.csv')

```

## Problem 1
**Provide a scatterplot of $LifeExp~TotExp$ and run simple linear regression. Do not transform the variables. Provide and interpret the F statistics, $R^2$, standard error, and p-values only. Discuss whether the assumptions of simple linear regression met.**

```{r}

plot(who.df$TotExp, who.df$LifeExp, main="Total Expenditure vs Life Expectancy", xlab="Total Expenditure", ylab="Life Expectancy")
```

```{r}

who.lm <- lm(LifeExp~TotExp, dat = who.df)
summary(who.lm)

```

* **F-statistic**: The F-test compares the current model to a model with one fewer predictor. If the current model is better than the reduced model, the p-value will be small. In this case, the p-value associated with the f-test is very small $(7.714e-14)$, so we can conclude that this model fits the data better than one with no independent variables. 
* $R^2$: This measures how well our model describes the data. The $R^2$ value of $0.2577$ means that the model explains only $25.77\%$ of the data’s variation.
* **Standard Error**:  This measures the variation in the residuals. As a general rule, this value should be 1.5 times the quartile values. For the first quartile: $4.778∗1.5 = 7.167$ and for the third quartile: $7.116∗1.5 = 10.674$. It is about 24% higher than the first quartile and 13% lower than the third quartile. 
* **P-Values**: This measures the probability that the factor is *not* important in the model. The smaller the value, the more significant the factor is. Both of the p-values are well below the standard threshold of $0.05$, so we know that the total expenditure is significant. 

**Are assumptions of linear regression met?** No -- we can see off the bat that the data does not seem to follow a linear trend (it looks logarithmic). Additionally, the variable that we've used only accounts for $25.77\%$ of the variance, so other factors must be at play. 


## Problem 2
**Raise life expectancy to the 4.6 power (ie. $LifeExp^{4.6}$). Raise total expenditure to the 0.06 power (nearly a log transform, $TotExp^{0.06}$). Plot $LifeExp^{4.6}$ as a function of $TotExp^{0.06}$ and re-run the simple regression model using the transformed variables. Provide and interpret the F statistics, $R^2$, standard error, and p-values. Which model is "better"?**

```{r}

who.df$LifeExp46 <- who.df$LifeExp^(4.6)
who.df$TotExp06 <- who.df$TotExp^(0.06)

plot(who.df$TotExp06, who.df$LifeExp46, main="Total Expenditure vs Life Expectancy - Transformed", xlab="Total Expenditure", ylab="Life Expectancy")
```


```{r}
who2.lm <- lm(LifeExp46~TotExp06, data = who.df)
summary(who2.lm)
  
```

* **F-statistic**: Once again, we see that the p-value associated with the F-statistic is very small (and it didn't change much from the last model), so we know that this model fits the data better than one with no independent variables. 
* $R^2$: The $R^2$ value has increased to $0.7298$, so we know that the model accounts for $72.98\%$ of the data’s variation. 
* **Standard Error**: The standard error is about $10.52\%$ higher than 1.5 times our first quartile $(53,978,977 * 1.5 = 80,968,466)$ and $1.97\%$ higher than 1.5 times the third quartile $(59,139,231 * 1.5 = 88,708,847)$. This is closer than our previous model.  
* **P-Values**: The p-value associated with our newly derived $TotExp06$ variable is much less than the standard threshold of $0.05$, so the variable is significant.

**Which model is better?** This model is much better because:

1. The plot looks more linear.
2. The $R^2$ value has increased drastically.
3. The standard error more closely matches 1.5 times the first and third quartiles.



## Problem 3
**Using the results from 2, forecast life expectancy when $TotExp^{0.06} = 1.5$. Then forecast life expectancy when $TotExp^{0.06}= 2.5$.**  
We know that our model predicts $LifeExp^{4.6}$, so in order to translate this to just $LifeExp$, we will need to do the following: $predictedValue^{\frac{5}{23}}$:

```{r}

predTotExp1 <- data.frame(TotExp06 = 1.5)
predTotExp2 <- data.frame(TotExp06 = 2.5)

pred1 <- predict(who2.lm, newdata=predTotExp1)
pred2 <- predict(who2.lm, newdata=predTotExp2)

paste0('For total expenditure = 1.5, the life expectancy is: ', pred1^(5/23))
paste0('For total expenditure = 2.5, the life expectancy is: ', pred2^(5/23))

```

## Problem 4
**Build the following multiple refression model and interpret the F statistic, $R^2$, standard error, and p-values. How good is the model?**  
$$LifeExp = b_0 + b_1PropMD + b_2TotExp + b_3(PropMD*TotExp) $$

```{r}

who.df$interaxn_term <- who.df$PropMD * who.df$TotExp

who3.lm <- lm(LifeExp ~ PropMD + TotExp + interaxn_term, data = who.df)
summary(who3.lm)


```

* **F-statistic**: The p-value associated with the F-statistic is very small (and it didn't change much from the last model), so we know that this model fits the data better than one with just 2 variables. 
* $R^2$: The $R^2$ value has decreased to $0.3574$, so we know that the model accounts for only $35.74\%$ of the data’s variation. 
* **Standard Error**: The standard error is about $29.29\%$ higher than 1.5 times our first quartile $(4.132 * 1.5 = 6.198)$ and $11.92\%$ lower than 1.5 times the third quartile $(6.540 * 1.5 = 9.81)$. This is worse than our previous model.  
* **P-Values**: The p-values associated with all three of the variables are much less than the standard threshold of $0.05$, so all variables are significant.

**How good is the model?** This model is not great at fitting the data because the $R^2$ value is low and the standard error is not around 1.5 times the first and third quartiles. Our previous model fit the data better. 


## Problem 5
**Forecast $LifeExp$ when $PropMD=0.3$ and $TotExp = 14$. Does this forecast seem realistic? Why or why not?**
```{r}

predPropMD <- 0.3
predTotExp3 <- 14
predInteraxn <- predPropMD * predTotExp3

predData <- data.frame(PropMD = predPropMD,
                       TotExp = predTotExp3,
                       interaxn_term = predInteraxn)


pred3 <- predict(who3.lm, newdata=predData)

pred3


```

**Does this forecast seem realistic? Why or why not?** No, this forecast does not seem realistic because an individual is not going to live to be 512 years old. 

