---
title: "DATA 605 - Assignment 9"
author: "Amber Ferger"
date: "3/27/2020"
output: html_document
---

## Problem 11 (pg.363)
**The price of one share of stock in the Pilsdoff Beer Company (see Exercise 8.2.12) is given by $Y_n$ on the $n$th day of the year. Finn observes that the differences $X_n = Y_{n+1} −Y_n$ appear to be independent random variables with a common distribution having mean $\mu = 0$ and variance $\sigma^2 = \frac{1}{4}$. If $Y_1 = 100$, estimate the probability that $Y_{365}$ is:**  

a. $\geq 100$  
b. $\geq 110$ 
c. $\geq 120$

By definition, we know that $S_n = X_1 + X_2 + X_3 + ... + X_n$. We are told that $X_n = Y_{n+1} - Y_n$, so subbing this in, we get: 
$$S_n = (Y_2 - Y_1) + (Y_3 - Y_2) + (Y_4 - Y_3)+ ... + (Y_{n+1} - Y_n)$$ 

Simplifying by canceling out terms, we are left with: 
$$S_n = Y_{n+1} - Y_1$$
Since we are asked to find $Y_{365}$, we will use $n=364$ to calculate $S_{364}=Y_{365}-Y_1$.   

Finally, we can substitute this into our equation: 
$$S_n^* = \frac{S_n - n\mu}{\sqrt{n} \sigma}$$

We'll define our common variables, $n$, $\mu$, $\sigma^2$, $\sigma$, and $Y_1$: 

```{r}
n <- 364 
mu <- 0
var <- 1/4
stdev <- sqrt(var)
y1 <- 100
```

### $Y_{365} \geq 100$ 
```{r}

y365 <- 100

s364 <- y365-y1

sn1 <- (s364 - (n*mu)) / (sqrt(n) * stdev)
pnorm(sn1, lower.tail=FALSE)


```

### $Y_{365} \geq 110$ 
```{r}

y365 <- 110

s364 <- y365-y1

sn2 <- (s364 - (n*mu)) / (sqrt(n) * stdev)
pnorm(sn2, lower.tail=FALSE)


```



### $Y_{365} \geq 120$ 
```{r}

y365 <- 120

s364 <- y365-y1

sn3 <- (s364 - (n*mu)) / (sqrt(n) * stdev)
pnorm(sn3, lower.tail=FALSE)


```

## Problem 2 
**Calculate the expected value and variance of the binomial distribution using the moment generating function.**  
The moment generating function is:  
$$M_x(t) = \Sigma e^{tx}p(x)$$
The binomial distribution is:  
$${ n \choose j}p^jq^{n-j}$$

Substituting this in, we get: 
$$M_x(t) = \Sigma e^{tx} \cdot { n \choose j}p^jq^{n-j}$$
This further evaluates to:  
$$M_x(t) = (pe^{t}+q)^n $$

We also know that:  
$$E(x) = \mu_1 = M_x'(0)$$
and $$\sigma^2 = \mu_2  - \mu_1^2 = M_x''(0) - M_x'(0)^2$$

### Expected Value
We know that the first derivative of the moment generating function evaluated at $t=0$ is the expected value, so we calculate this as: 
$$M'_x(t) = n(pe^t+q)^{n-1}\cdot pe^t$$
Evaluated at $t=0$: 
$$E(x)= \mu_1 = npe^0(pe^0+q)^{n-1}$$
$$= np(p+q)^{n-1}$$
We know that $p+q=1$, so: 
$$ E(x)= \mu_1 =  np$$

### Variance 
By definition, we know that the variance is:  
$$\sigma^2 = \mu_2 − \mu_1^2$$

Let's first calculate $M''_x(t)$:   
$$M''_x(t) = \mu_2 = [npe^t(pe^t+q)^{n-1}]+[npe^t \cdot (n-1)(pe^t+q)^{n-2} \cdot pe^t]$$

Evaluated at $t=0$:  

$$= [npe^0(pe^0+q)^{n-1}]+[npe^0 \cdot (n-1)(pe^0+q)^{n-2} \cdot pe^0]$$
$$= [np(p+q)^{n-1}]+[np \cdot (n-1)(p+q)^{n-2} \cdot p]$$
Once again, we know that $p+q=1$, so this simplifies to: 
$$\mu_2 = np+ np^2(n-1)$$
Finally, subbing in for our variance equation $\sigma^2 = \mu_2 − \mu_1^2$:  
$$\sigma^2 = np+ np^2(n-1) - (np)^2$$
$$= np+ n^2p^2 - np^2 - n^2p^2$$

$$\sigma^2= np- np^2$$

## Problem 3 
**Calculate the expected value and variance of the exponential distribution using the moment generating function.**  
We will start out by creating the moment function:  
$$M_x(t) = \int_0^\infty e^{tx}(\lambda e^{-\lambda x})dx$$ 

$$= \lambda \int_0^\infty e^{x(t -\lambda)}dx$$ 
When $\lambda < t$: 
$$= \lambda [\frac{1}{t-\lambda}e^{-\infty}-\frac{1}{t-\lambda}e^{0}]$$

$$M_x(t)= \frac{-\lambda}{t-\lambda} = \frac{\lambda}{\lambda - t} $$


### Expected Value
$$M_x'(t) = \frac{0( \lambda - t) - \lambda(-1)}{(\lambda - t)(\lambda - t)}$$ 
$$= \frac{\lambda}{\lambda^2 - 2 \lambda t + t^2}$$ 

Evaluated at $t=0$: 
$$E(x) = \mu_1 = \frac{\lambda}{\lambda^2} = \frac{1}{\lambda}$$ 

### Variance 
$$M_x''(t) =\mu_2 = \frac{0(\lambda^2 - 2 \lambda t + t^2) - \lambda (-2 \lambda + 2t)}{(\lambda^2 - 2 \lambda t + t^2)(\lambda^2 - 2 \lambda t + t^2)}$$

$$ = \frac{ 2\lambda^2 - 2 \lambda t}{\lambda^4 - 2 \lambda^3 t + \lambda^2t^2 - 2 \lambda^3 t + 4\lambda^2 t^2 - 2\lambda t^3 + t^2 \lambda^2 - 2 \lambda t^3 + t^4}$$ 

$$ = \frac{ 2\lambda(\lambda - t)}{\lambda^4 + t^4 - 4 \lambda^3 t + 6\lambda^2t^2 - 4 \lambda t^3}$$

Evaluated at $t=0$: 
$$ = \frac{ 2\lambda^2}{\lambda^4} = \frac{2}{\lambda^2}$$

Finally, subbing in for our variance equation $\sigma^2 = \mu_2 − \mu_1^2$: 
$$\sigma^2 = \frac{2}{\lambda^2} - (\frac{1}{\lambda})^2$$

$$= \frac{2}{\lambda^2} - \frac{1}{\lambda^2} $$

$$\sigma^2 = \frac{1}{\lambda^2}$$